---
title: "Lab One"
author: "Jalen Jiang"
date: "4/12/2020"
output:   
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Version

Here is the R code that will show the version of R I have installed, and below it is my output:
```{r}
version
```

## Loading Library

I will now load some libraries

```{r}
library(sf)
library(raster)
library(spData)
library(spDataLarge)
```

## Example 1: Map of Africa and South America

Here is some code that will plot a map of countries in Africa and South America, but no others. We start with `world` which is the set of all country multipolygons in the world, and create a filter for the column `world$continent` for Africa and South America. Then we plot column 0, the geometries.

```{r world}
world_asia = world[world$continent=="Africa" | world$continent == "South America", ]
plot(world_asia[0])
```

## Example 2: 10x10 raster

Here we create a 10x10 raster and fill them each with numbers ranging from 0 to 10. Which, in some sense, actually makes it 10x10x10...

```{r}
hundred_random_numbers = runif(100,0,10)
my_raster = raster(nrows = 10, ncols = 10, res = 1, xmn = 0, xmx = 10, ymn = 0, ymx = 10, vals = hundred_random_numbers)
plot(my_raster)
```